---
title: "Exam 3"
Author: "Alyssa Tidwell"
output: github_document
---

```{r setup, include=FALSE}
library(rmarkdown)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(modelr)
library(lattice)
```
#### YAY for ecology!  
  **First, what does our data look like?**   
```{r}
BioLogData <- read.csv(file = "./BioLogData_Exam3.csv", sep = "|")
head(BioLogData) 


```

Is it all in the ideal class type?
```{r}
sapply(BioLogData, class)

```
Hm. Looks alright to me.  
Let's start to summarize and visualize the paired data.

```{r}
summary(BioLogData)
pairs(BioLogData)
```  

### Question 1:  
**Which sample locations are functionally different from each other in terms of what C-substrates they can utilize?**    
The sample locations are:
```{r, echo=FALSE}
levels(BioLogData$Sample.ID)
```
And how many carbon sources are there?
```{r, echo=FALSE}
nlevels(BioLogData$Substrate)
```
Let's graph subsets of each location to help us visualize the how the carbon sources are used.   

```{r, echo=FALSE, eval=TRUE, results='hide'}
creek <- subset(x = BioLogData, Sample.ID == "Clear_Creek")
as.data.frame(creek)
creek <- pivot_longer(data = creek, cols= c(Hr_144, Hr_48, Hr_24), 
                      names_to = "inc_hrs", values_to = "values") 

soil1 <- subset(x = BioLogData, Sample.ID == "Soil_1")
as.data.frame(soil1)
soil1 <- pivot_longer(data = soil1, cols= c(Hr_144, Hr_48, Hr_24), 
                      names_to = "inc_hrs", values_to = "values") 

soil2 <- subset(x = BioLogData, Sample.ID == "Soil_2")
as.data.frame(soil2)
soil2 <- pivot_longer(data = soil2, cols= c(Hr_144, Hr_48, Hr_24), 
                      names_to = "inc_hrs", values_to = "values") 


wastewater <- subset(x = BioLogData, Sample.ID == "Waste_Water")
as.data.frame(wastewater)
wastewater <- pivot_longer(data = wastewater, cols= c(Hr_144, Hr_48, Hr_24),names_to = "inc_hrs", values_to = "values") 


```
```{r, echo=FALSE, eval=TRUE}

ggplot(data=creek, aes(x = Substrate, y= values, color=inc_hrs)) +
  geom_point() + geom_smooth(method = "lm")+
  labs(title= "Clear Creek Values", subtitle= "Absorbancy rates after 24, 48, and 144 hours of incubation",
       x="Carbon Source", y="Light absorbance value")+
  scale_color_discrete(name = "Incubation time")+   
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  theme(plot.background = element_rect(colour = "black"))

ggplot(data=soil1, aes(x = Substrate, y= values, color=inc_hrs)) +
  geom_point() + geom_smooth(method = "lm")+
  labs(title= "Soil 1 Values", subtitle= "Absorbancy rates after 24, 48, and 144 hours of incubation" , x="Carbon Source", y="Light absorbance value")+
  scale_color_discrete(name = "Incubation time")+  
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  theme(plot.background = element_rect(colour = "black"))

ggplot(data=soil2, aes(x = Substrate, y= values, color=inc_hrs)) +
  geom_point() + geom_smooth(method = "lm")+
  labs(title= "Soil 2 Values", subtitle= "Absorbancy rates after 24, 48, and 144 hours of incubation",
       x="Carbon Source", y="Light absorbance value")+
  scale_color_discrete(name = "Incubation time")+
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  theme(plot.background = element_rect(colour = "black"))

ggplot(data=wastewater, aes(x = Substrate, y= values, color=inc_hrs)) +
  geom_point() + geom_smooth(method = "lm")+
  labs(title= "Waste Water Values", subtitle= "Absorbancy rates after 24, 48, and 144 hours of incubation",
       x="Carbon Source", y="Light absorbance value")+
  scale_color_discrete(name = "Incubation time")+  
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  theme(plot.background = element_rect(colour = "black"))

```  
  
There is variance in the total amount of cabon being used and the substrate that is used as a carbon source. It looks like the soil samples are functionally different than the water sources because the water samples have less diversity in carbon sources used and lower values overall.  
      
### Question 2:  
**Are Soil and Water samples significantly different overall? What about for individual carbon substrates?**     

Let's make a new column of data, "type", that will allow us to see whether a sample is water-sourced or soil-sourced. We'll model with our subsets to see if there's a significant difference between the types, and simutaneously gain information about substrate significance. Because some of our variables are categorical and we later want to run a Tukey's test, we'll start with a linear fit but ultimately want to use AOV for these models. 


```{r, echo=FALSE, eval=TRUE}

creek <- creek %>% 
  mutate(type="water")
wastewater <- wastewater %>% 
  mutate(type="water")
soil1 <- soil1 %>%
  mutate(type="soil")
soil2 <- soil2 %>%
  mutate(type="soil")

bound <- rbind(creek, wastewater, soil1, soil2)

```
```{r, echo=TRUE, eval=TRUE}
moda <- aov(data = bound, values ~ Substrate * type)
summary(moda)
modb <- aov(data = bound, values ~ Substrate+type)
summary(modb)

```
  According to these models, the sample type is a predictor of values, and so is carbon source.  Let's try to visualize the difference of utilization for each carbon source.
```{r, echo=FALSE, eval=TRUE}

ggplot(data=bound, aes(x=type, y=values, color = type))+
  geom_boxplot()+
  facet_wrap(~Substrate)

```

  
### Question 3:  
**If there are differences between samples, which C-substrates are driving those differences?**  
The C-substrates that are driving the differences are any substrate with the higheset significance codes from the following list:
```{r, echo=FALSE , eval= TRUE}
modc <- lm(data = bound, values ~Substrate*type)
summary(modc)

```

### Question 4:  
**Does the dilution factor change any of these answers?**    
```{r, echo=FALSE, eval=TRUE}

mod1 <- aov(data=creek, creek$values~creek$Dilution)
plot(creek$values~creek$Dilution)
abline(mod1)
summary(mod1)


mod2 <- aov(data=soil1, soil1$values~soil1$Dilution)
plot(soil1$values~soil1$Dilution)
abline(mod2)
summary(mod2)


mod3 <- aov(data=soil2, soil2$values~soil2$Dilution)
plot(soil2$values~soil2$Dilution)
abline(mod3)
summary(mod3)

mod4 <- aov(data=wastewater, wastewater$values~wastewater$Dilution)
plot(wastewater$values~wastewater$Dilution)
abline(mod4)
summary(mod4)


```
  
Less carbon is being consumed as the concentration of the soil samples increases.  
More carbon is being consumed as the concentration of the water samples increases.
  
### Question 5:  
**Do the control samples indicate any contamination?**    
Water is the negative control. If the BioLog machine reads a light absorbtion value other than 0 for the negative control, there is contamination.
```{r, echo=FALSE, eval=TRUE}
water <- BioLogData %>%
  filter(BioLogData$Substrate == "Water")
water
```
Looking at the values under Hr_24, Hr_48, and Hr_144, we can see that all of the values are 0, telling us that there was no contamination measured at any point during incubation.